---
title: "Lab 2"
author: "Kristen Fitzgerald"
subtitle: "AA 501"
date: '2024-06-26'
output: pdf_document
header-includes:
  - \usepackage{setspace}
  - \doublespacing
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(UsingR)
library(ggplot2)
library(dplyr)
data(normtemp)
data(AirPassengers) 
```

# Question 1

Revisit the NormTemp dataset from Lab 1, where we examined the observed
mean body temperature (`temperature`) in comparison to the well-known
“average” of 98.6.

__a.__ Perform a statistical test (alpha = 0.05) to determine whether this 
well known number is actually the mean body temperature. What is your
p-value? Explain in words what this p-value means. What is your
conclusion?

__Solution:__ First, we check normality and find that `temperature` is normally
distributed. We define our hypotheses to be $H_{0}: \mu = 98.6$ vs. 
$H_{A}: \mu \ne 98.6$. We find a p-value of less than 0.001. This means that 
there is essentially a zero percent chance of observing a mean temperature 98.6
using this data. Thus, we reject the null hypothesis and conclude that there is 
statistically significant evidence to conclude that the mean mean body 
temperature differs from 98.6.  

\singlespacing

```{r Q1a.norm}
ggplot(normtemp, aes(sample = temperature)) +
  stat_qq() + 
  stat_qq_line()
```


```{r Q1a.test}
t.test(normtemp$temperature, mu = 98.6)
```

\doublespacing

\newpage

__b.__ Give the 95% Confidence Interval for temperature. Explain in words what
a 95% confidence interval represents.

__Solution:__ We are 95% confident that the true mean body temperature falls
between 98.12200 and 98.37646. 

\bigskip

__c.__ If we restrict our analysis to only the females in this dataset, would our
conclusion change? 1 = male, 2 = female

__Solution:__ No, our conclusion does not change. There is statistically
significant evidence to conclude that the average female body differs from 98.6.

\singlespacing

```{r 2c}
normtemp_female <- normtemp %>%
  filter(gender == 2)

t.test(normtemp_female$temperature, mu = 98.6)
```

\doublespacing

\bigskip

__d.__ Is there any difference (alpha=0.05) in temperature between the two
genders recorded in this dataset (be sure to look at assumptions
and perform the correct test)?

__Solution:__ First, we check normality and conclude that both genders 
approximately follow a normal distribution. Next, we check if temperature has
equal variance across both groups. We find that there is equal variance across
groups. We perform a two sample t-test and find a p-value of 0.02393. Thus,
we conclude that there is a statistically significant difference between average
body temperature in males and females.

\singlespacing

```{r Q1D.Assumptions}
shapiro.test(normtemp$temperature[normtemp$gender == "1"])
shapiro.test(normtemp$temperature[normtemp$gender == "2"])

var.test(temperature ~ gender, data = normtemp) 
```

```{r Q1D.test}
t.test(temperature ~ gender, data = normtemp, var.equal = TRUE)
```

\doublespacing

\newpage

# Question 2

The Airline dataset contains information regarding the number of international
airline travelers (variable air) across different months of the year from
1949-1960. 

__a.__ We are interested in knowing if during this time period there was a
significant difference between air travel in the Summer months of June,
July, and August vs. the remainder of the year? Use a statistical
hypothesis test (alpha=0.05) to support your answer.

__Solution:__ The normality assumption fails for non-summer months. We will use
a Wilcoxin rank test. We find that the difference between air travel in the
summer and non-summer months is significantly different. 

```{r Q2, include=F}
library(tseries)
library(forecast)
cycle(AirPassengers)

air1 = data.frame(AirPassengers)
air2 = air1 %>% mutate(summer=ifelse(cycle(AirPassengers) %in% 6:8,1,0))
```

\singlespacing

```{r Q2.Assumptions}
shapiro.test(air2$AirPassengers[air2$summer == "1"])
shapiro.test(air2$AirPassengers[air2$summer == "0"])
```

```{r Q2.test}
wilcox.test(AirPassengers ~ summer, data=air2)
```

