---
title: "Lab 1"
author: "Kristen Fitzgerald"
date: '2024-06-25'
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# install.packages('UsingR')
# install.packages("AmesHousing")
library(AmesHousing)
library(UsingR)
library(ggplot2)
data(normtemp)
ames <- make_ordinal_ames()
```

# Question 1

Use the normtemp dataset to answer the following:

__a.__ Determine the following statistics for the variable temperature: Min, Max
mean, and standard deviation

```{r Q1a}
summary(normtemp$temperature)
```

\newpage

__b.__ Does temperature appear to be normally distributed?

__Solution:__ Yes, temperature does appear to be normally distributed.

```{r Q1b}
ggplot(normtemp) +
  geom_histogram(aes(x = temperature), bins = sqrt(130)) +
  labs(x = "Temperature")


ggplot(normtemp, aes(sample = temperature)) +
  stat_qq() +
  stat_qq_line()
```

\newpage

__c.__ Create box plots for temperature. Are there any outliers? Display 
a reference line at 98.6.

+ For horizontal line: + geom_hline(yintercept=98.6)

+ For vertical line: + geom_vline(xintercept=98.6)

Does the median body temperature seem to be 98.6 degrees?

__Solution:__ Yes, there are three outliers. No, the median is less than 98.6.

```{r Q1c}
ggplot(normtemp, aes(x = temperature)) +
  geom_boxplot() +
  geom_vline(xintercept=98.6)
```

\newpage

# Question 2

Using the Ameshousing dataset from our in-class examples, run some
distributional analysis on Sale_Price, Log(Sale_Price), and Gr_Liv_Area.

__a.__ Create histograms of these three variables.

+ Overlay a kernel density estimator of the variables.

```{r Q2a}
ggplot(ames, aes(x = Sale_Price)) +
  geom_histogram(aes(y=..density..), alpha=0.5, bins = sqrt(2930)) +
  labs(x = "Sale Price") +
  geom_density(alpha=0.5, col="red")
  
ggplot(ames, aes(x = log(Sale_Price))) +
  geom_histogram(aes(y=..density..), alpha=0.5, bins = sqrt(2930)) +
  labs(x = "Logarithm of Sale Price") +
  geom_density(alpha=0.5, col="red")

ggplot(ames, aes(x = Gr_Liv_Area)) +
  geom_histogram(aes(y=..density..), alpha=0.5, bins = sqrt(2930)) +
  labs(x = "Ground Living Area") +
  geom_density(alpha=0.5, col="red")
```

\newpage

__b.__ Create a QQ Plot for both Sale_Price and Log(Sale_Price). Based on
these exploratory procedures, which version of the price information
(Sale_Price or Log(Sale_Price)) would you say is closer to being normally
distributed?

__Solution:__ The logarithm of `Sale_Price` is more normally distributed.

```{r Q2b}
ggplot(ames, aes(sample = Sale_Price)) +
  stat_qq() +
  geom_qq_line()

ggplot(ames, aes(sample = log(Sale_Price))) +
  stat_qq() +
  geom_qq_line()
```

\newpage

__3.__ Using the Ameshousing dataset from our in-class examples, determine the
following:

__a.__ What type of variables are each of these columns (Nominal, Ordinal, or
Continuous/Quantitative)? Keep in mind that the way they are
represented in the R dataset may not be appropriate, so you should make
this determination using your own judgement based on the data you are
looking at.

- __Overall_Qual__
    - Represented as Ordinal
    - Should be Ordinal

- __Lot_Shape__
    - Represented as Ordinal
    - Should be Ordinal

- __Heating_QC__
    - Represented as Ordinal
    - Should be Ordinal

- __Lot_Area__
    - Represented as numeric
    - Should be numeric

```{r Q3}
class(ames$Overall_Qual)
class(ames$Lot_Shape)
class(ames$Heating_QC)
class(ames$Lot_Area)
```

