---
title: "Lab 3"
subtitle: "AA 501"
author: "Kristen Fitzgerald"
date: '2024-06-27'
output: pdf_document
header-includes:
  - \usepackage{setspace}
  - \doublespacing
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
garlic <- read.csv("https://raw.githubusercontent.com/IAA-Faculty/statistical_foundations/master/garlic.csv")
bottle <- read.csv("https://raw.githubusercontent.com/IAA-Faculty/statistical_foundations/master/bottle.csv")
trials <- read.csv("https://raw.githubusercontent.com/IAA-Faculty/statistical_foundations/master/trials.csv")
garlic$Fertilizer <- factor(garlic$Fertilizer)
bottle$Line <- factor(bottle$Line)
library(ggpubr)
library(ggplot2)
library(car)
library(stats)
library(DescTools)
library(dunn.test)
```

# Question 1 

Use the dataset Garlic to perform an ANOVA. This dataset has information on
the weight of garlic bulbs (`bulbwt`) for 32 different garlic plants. Each
garlic plant was initially treated with 1 of 4 types of `fertilizer.` The
purpose of the experiment is to determine whether or not there is any
difference in the resulting bulb weight when a garlic plant is treated with
a different fertilizer.

__a.__ Test the hypothesis that the means of bulb weight are equal
regardless of the fertilizer used.

+ Verify that the assumption of normality in each sample you are
testing. Are you comfortable with this assumption?

+ Verify that the groups satisfy the assumption of equal variance.

+ State your conclusion to the hypothesis test (alpha = 0.05)

__Solution:__ The data satisfies the normality & equal variance assumption.
We reject the null hypothesis, there is significant evidence to conclude that 
the average bulb weight differs depending on the fertilizer used.

## Verifying Normality

\singlespacing

```{r 1a.Norm}
model <- aov(BulbWt ~ Fertilizer, data=garlic)

ggqqplot(residuals(model))

shapiro.test(residuals(model))
```

\doublespacing

\newpage

## Veryifying Equal Variances

```{r 1a.Var}
leveneTest(BulbWt ~ Fertilizer, data=garlic)

ggplot(model, aes(.fitted, .resid)) + 
  geom_point() +
  geom_hline(yintercept = 0)+
  labs(title = 'Residual vs. Fitted Values Plot',
       x = 'Fitted Values', y = 'Residuals')
```

\newpage

## Basic ANOVA

```{r 1a.ANOVA}
model<-aov(BulbWt ~ Fertilizer, data=garlic)

summary(model)
```

\doublespacing

\newpage

__b.__ If you were to test if each type of fertilizer were different from each
other fertilizer (pairwise), how many hypothesis tests would
you be running?

__Solution:__ With for groups to compare you would need six hypothesis tests.

__c.__ If the probability of incorrectly rejecting a true null hypothesis is
0.05, and each hypothesis test is considered independent,
what is the probability that you incorrectly reject at least one
true hypothesis? How do we solve this problem when we’re
performing post-hoc analysis in ANOVA?

__Solution:__ The experimentwise error rate at a 0.05 level would be 0.26.
We solve this problem by using pairwise or control group comparison methods.

__d.__ Which fertilizers are statistically different from each other and which
fertilizers do not appear to produce different bulb weights?

__Solution:__ Fertilizer 4 differs significantly from groups 1 and 3.

\singlespacing

```{r 1d.Tukey}
tukey.garlic <- TukeyHSD(model)
print(tukey.garlic)
par(mar=c(4,10,4,2))
plot(tukey.garlic, las=1)
```

\newpage

# Question 2

The Bottle dataset contains observations from a factory that is producing
plastic water bottles along 3 different assembly lines.The number of units 
produced by each assembly line are given for a
number of days. The manufacturer wants to know if the assembly lines are
producing the same number of bottles. He suspects that 2 of his lines are
significantly better than the third, but he has had trouble demonstrating this
due to the variability in production. Confirm or deny his hypothesis using
the data provided. Be sure to verify your assumptions before proceeding
with the analysis.

__Solution:__ Group 3 differs significantly from Groups 1 and 2. Thus,
confirming his hypothesis that two lines are performing better. 

## Verifying Normality

\singlespacing

```{r 2a.Norm}
model <- aov(Units ~ Line, data=bottle)

ggqqplot(residuals(model))

shapiro.test(residuals(model))
```

\newpage

## Veryifying Equal Variances

```{r 2a.Var}
leveneTest(Units ~ Line, data=bottle)

ggplot(model, aes(.fitted, .resid)) + 
  geom_point() +
  geom_hline(yintercept = 0)+
  labs(title = 'Residual vs. Fitted Values Plot',
       x = 'Fitted Values', y = 'Residuals')
```

\newpage

## Basic ANOVA

```{r 2a.ANOVA}
model<-aov(Units ~ Line, data=bottle)

summary(model)
```

## Post Hoc Test

```{r 2d.Tukey}
tukey.bottle <- TukeyHSD(model)
print(tukey.bottle)
par(mar=c(4,10,4,2))
plot(tukey.bottle, las=1)
```

\doublespacing

# Question 3

The `Trials` dataset contains information from a clinical trial for blood pressure
medicine. There are two different drugs being compared to a placebo, 1 of which 
is already FDA approved. The study aims to confirm that the new drug
(treatment = New Drug) is effective at lowering blood pressure when
compared to the placebo. Perform an ANOVA and post-hoc analysis which
compares the reduction in blood-pressure (`BPChange`) of the new drug
and the existing drug (treatment = Approved Drug) to the placebo. Do
both drugs outperform the placebo in blood-pressure reduction? Be sure
to verify your assumptions before proceeding with the analysis.

## Verifying Normality

\singlespacing

```{r 3.Norm}
model <- aov(BPChange ~ Treatment, data=trials)

ggqqplot(residuals(model))

shapiro.test(residuals(model))
```

\newpage

## Veryifying Equal Variances

```{r 3.Var}
leveneTest(BPChange ~ Treatment, data=trials)

ggplot(model, aes(.fitted, .resid)) + 
  geom_point() +
  geom_hline(yintercept = 0)+
  labs(title = 'Residual vs. Fitted Values Plot',
       x = 'Fitted Values', y = 'Residuals')
```


\newpage

## Welch's ANOVA

```{r 3.ANOVA}
oneway.test(BPChange ~ Treatment, data = trials, var.equal = FALSE)
```

## Post-Hoc Test

```{r 3.PostHoc}
trials$Treatment <- ifelse(trials$Treatment=="Placebo","aPlacebo", trials$Treatment)
DunnettTest(x = trials$BPChange, g = trials$Treatment)
```

